---
title: "OrthoFinder"
author: "Emily Giroux"
date: "5/28/2018"
output: html_document
---

**Getting started in R:** 
Set the working directory > setwd("~/") Check version installed
```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy = TRUE, fig.align='center')
```

**This will help us when finding our files to source functions:**
```{r sourcing_my_functions}
install.packages("rprojroot")
library(rprojroot)
# We specify ours is an RStudio project
# The root object contains a function that will help us locate our package r files regardless of our current working directory
root <- rprojroot::is_rstudio_project
scriptsPath <- root$make_fix_file(".")("R")
scripts  <- dir(root$find_file("R", path = root$find_file()))
scriptsl <- paste(scriptsPath, scripts, sep = "//")
lapply(scriptsl, source)
```

**Packages to install and have ready for later analyses:**
```{r}
install.packages("data.table")
```

**User: Define the path to the shared folder where the main working directory will be.**
```{r setting_the_main_directory, cache=TRUE}
sharedPath <- "/isilon/cfia-ottawa-fallowfield/users/girouxeml/PIRL_working_directory/"
```

**Define where the analysis path is, or the path to the current project:**
```{r setting_the_main_directory, cache=TRUE}
analysisOrtho     <- "Lachnellula_species_GenomeAn_IonTorrent_2017/"
sharedPathAnOrtho <- paste(sharedPath, analysisOrtho, sep="")
```

**Define path variables to programs and scripts used:**
```{r}
# Biocluster system-wide programs:
bedtoolsPath    <- "/opt/bio/BEDTools/bin/bedtools"
blastallPath    <- "/opt/bio/ncbi/bin/blastall"

# *** Revisit this organization. Perhaps best that all programs be /home/ and databases be in local cfia-ottawa?
# CFIA-ACIA users home directory programs:
progPath        <- "/home/CFIA-ACIA/girouxeml/prog/"
blastpPath      <- paste(progPath, "miniconda/bin/blastp", sep = "")
buscoPath       <- paste(progPath, "busco/scripts/run_BUSCO.py", sep = "")
jbrowseFlatfile2json <- paste(progPath, "jbrowse/JBrowse-1.12.3/bin/flatfile-to-json.pl", sep = "")
jbrowsePrepRefsSeqsPath <- paste(progPath, "jbrowse/JBrowse-1.12.3/bin/prepare-refseqs.pl", sep = "") 
scriptsPath     <- paste(progPath, "scripts_pl/", sep = "")
orthoFinderPath <- paste(progPath, "OrthoFinder-2.2.6/orthofinder", sep = "")
# CFIA-Ottawa-Fallowfield user directory programs:
programsPath    <- "/isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/"
```

**Define paths to reference files used:**  
```{r}
referencesPath  <- paste(sharedPath, "References/", sep = "")
cdnaBotcinPath  <- paste(referencesPath, "Botrytis_cinerea.ASM83294v1.cdna.all.fa", sep = "")
gff3BotcinPath  <- paste(referencesPath, "Botrytis_cinerea.ASM83294v1.37.gff3", sep = "")
pepBotcinPath   <- paste(referencesPath, "Botrytis_cinerea.ASM83294v1.pep.all.fa", sep = "")
pepBotcinT4Path <- paste(referencesPath, "Botrytis_cinerea_t4.BotFuc_Mar2011.pep.all.fa", sep = "")
pepGlarlozPath  <- paste(referencesPath, "Glarea_lozoyensis_atcc_20868.GLAREA.pep.all.fa", sep = "")
pepMarbrunPath  <- paste(referencesPath, "Marsonnina_brunnea_GCA_000298775.1_ASM29877v1_protein.faa", sep = "")
pepPhiascoPath  <- paste(referencesPath, "Phialocephala_scopiformis.Phisc1.pep.all.fa", sep = "")
pepPhiasubPath  <- paste(referencesPath, "Phialocephala_subalpina.PAC_version_1.pep.all.fa", sep = "")
pepRhyncomPath  <- paste(referencesPath, "Rhynchosporium_commune.version_1.pep.all.fa", sep = "")
pepScleborPath  <- paste(referencesPath, "Sclerotinia_borealis_f_4128.SBOR_1.pep.all.fa", sep = "")
pepSclerscPath  <- paste(referencesPath, "Scerotinia_sclerotiorum_GCF_000146945.2_ASM14694v2_protein.faa", sep = "")
```

Start orthoFinder from scratch on a set of protein fasta files from different species, saved to a directory:
```{r}
prefix <- "Test2_OrthoFinder"

orthoAnalysisPath <- paste(sharedPathAnOrtho, "orthofinderAnalysis/", sep = "")
pathProtFasOrtho  <- paste(orthoAnalysisPath, "proteinFasta_descriptive/protein", sep = "")
node <- 20
threadsAnalysis <- 20
nameResultsDir <- "test1"

cmd <- paste("cd ", orthoAnalysisPath, " && ", orthoFinderPath,
             " -t ", node, " -a ", threadsAnalysis, " -M msa ",
             " -n ", nameResultsDir, 
             " -f ", pathProtFasOrtho, sep = "")
cmd
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node) 
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Start OrthoFinder from a later point, after an analysis has been done and Blast databases already made:
```{r}
prefix <- "Test1_OrthoFinderMSATrees"
node   <- 20
threadsAnalysis <- 20
previousResultsDir  <- "Results_test1_May24"
previousResultsPath <- paste(pathProtFasOrtho, previousResultsDir, sep = "")

cmd <- paste("cd ", previousResultsPath, " && ", orthoFinderPath,
             " -t ", node, " -a ", threadsAnalysis, " -M msa -fg . ", sep = "")
cmd
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node) 
```
For some reason, specifying nodes in the qsub doesn't end up using more than 1 node, despite directing it to do so.  
However, running this directly in the directory with the bash files does:  

qsub -pe smp 20 -cwd -S /bin/bash "name of bash file"  
  
Note that the temp .e and .o files will be written to the home directory and will need to be cleaned up from there.


Read inthe csv file that has the single copy orthogroups list with gene names:
```{r}
orthoList <- paste(pathProtFasOrtho, "Results_test1_May24/orthoFinder_singleCopy_subset.csv", sep = "")
orthoListTbl <- fread(file = orthoList, sep = "auto", header = FALSE, data.table = TRUE)
```

Will need to retrieve the genes, not the proteins. For marker development we will be testing on DNA, not RNA, and primers
designed across intron/exon boundaries need to be avoided. 

Using Hai's orthofinder pipeline. Use genbank_to_fasta on the genbank .gbf files:  
$ ~/prog/genbank_to_fasta/genbank_to_fasta_v1.2/genbank_to_fasta.py -i Lari.gbf -m genbank -s nt -f CDS -d pipe -q '"ncbi","Lachnellula_arida",product,locus_tag'  
$ sed -i 's/missing_"ncbi"_qualifer/ncbi/g' L*.fasta  
$ sed -i 's/ /_/g' L*.fasta  
$ sed -i 's/missing_"//g' L*.fasta  
$ sed -i 's/"_qualifer//g' L*.fasta  

$ for file in ./L*.gbf  
> do 
> ~/prog/genbank_to_fasta/genbank_to_fasta_v1.2/genbank_to_fasta.py -i "$file" -m genbank -s nt -f CDS -d pipe -q 'locus_tag'
> done





